Economic Theory of Fertility Explains Modern Natural Selection

### 2020-11-19 why-negative-selection.Rmd
remove partners-controlled regression until you can figure out how to do it right

We next focus on variables related to mate search and household
formation. We split males and females by lifetime number of sexual
partners, at the median value of 3 (Figure \@ref(fig:plot-n-partners)a).
For both sexes, selection effects are larger and more significant among
those with more than 3 partners. Results are unchanged if we control for
respondents' age (appendix Figure
\@ref(fig:plot-n-partners-controlled)). Splitting respondents by whether
they were living with a spouse or partner at the time of interview,
effects are larger among those not living with a spouse or partner
(Figure \@ref(fig:plot-n-partners)b). Lastly, we split female
respondents by age at first live birth (AFLB).[^1] There is evidence for
genetic effects on AFLB [@Barban_2016], and there is a close link
between this variable and number of children born. Figure
\@ref(fig:plot-n-partners)c shows effect sizes estimated separately for
each tercile of AFLB. Several effects are strikingly different across
terciles. ADHD and MDD are selected for amongst the youngest third of
mothers, but selected against among the oldest two-thirds. Educational
attainment is selected for among the oldest two-thirds of mothers, but
is not significantly selected among the youngest third. Similarly,
several PGS of body measurements are selected against only among older
mothers. The correlation between effect sizes for the youngest and
oldest terciles is `r cor_age_flb`. As before, recall that PGS may have
two effects: changing the number of children within each category, and changing
the number of children by shifting respondents between categories.

@@

## Selection by number of partners, controlling for age

Figure \@ref(fig:plot-n-partners-controlled) shows selection effects by
lifetime number of sexual partners. Age, age squared, and interactions
of age and age squared with PGS are controlled for.

```{r plot-n-partners-controlled, fig.cap = "Selection effects by number of sexual partners (median split at 3), controlling for respondent's age and age squared."}
drake::loadd(res_partners_controlled)
res_partners_controlled %<>%       
      mutate(`N partners` = ifelse(grepl("TRUE", term), "<= 3", "> 3"))

n_regs <- as.double(nrow(res_partners_controlled))
res_partners_controlled %>% 
      mutate(
        score_name = pretty_names(score_name)
      ) %>% 
      rename(p = p.value) %>% 
      ggplot(aes(
        y      = reorder_within(score_name, estimate, sex, fun = order_abs(1)), 
        x      = estimate, 
        fill   = `N partners`, 
        colour = p < 0.05/{{n_regs}}
      )) + 
        my_vline +
        pval_geom_point() +
        facet_wrap(vars(sex), scales = "free_y") +
        my_fill_scale(n = 2) +
        pval_scale() +
        my_theme(legend.position = "bottom") +
        my_labs(x = "", y = "") + 
        scale_y_reordered()

>>>>

We next focus on variables related to mate search and household
formation. We split males and females by lifetime number of sexual
partners, at the median value of 3 (Figure \@ref(fig:plot-n-partners)a).
For both sexes, selection effects are larger and more significant among
those with more than 3 partners. Splitting respondents by whether
they were living with a spouse or partner at the time of interview,
effects are larger among those not living with a spouse or partner
(Figure \@ref(fig:plot-n-partners)b). Lastly, we split female
respondents by age at first live birth (AFLB).[^1] There is evidence for
genetic effects on AFLB [@Barban_2016], and there is a close link
between this variable and number of children born. Figure
\@ref(fig:plot-n-partners)c shows effect sizes estimated separately for
each tercile of AFLB. Several effects are strikingly different across
terciles. ADHD and MDD are selected for amongst the youngest third of
mothers, but selected against among the oldest two-thirds. Educational
attainment is selected for among the oldest two-thirds of mothers, but
is not significantly selected among the youngest third. Similarly,
several PGS of body measurements are selected against only among older
mothers. The correlation between effect sizes for the youngest and
oldest terciles is `r cor_age_flb`. As before, recall that PGS may have
two effects: changing the number of children within each category, and changing
the number of children by shifting respondents between categories.

@@
